Invariant Representations for Noisy Speech Recognition

### 2016-10-24 paper.tex
Added all references

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel and recording conditions. Ensuring such robustness to variability is a challenge in modern day neural network-based ASR systems, especially when all types of variability are not seen during training. We attempt to address this problem by encouraging the neural network acoustic model to learn invariant feature representations.
    We use ideas from recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training. A recent work from Ganin et al. proposes to
    use adversarial training for image domain adaptation by using an intermediate
    representation from the main target classification network to deteriorate the domain 
    classifier performance through a separate neural network.
    Our work focuses on investigating neural architectures which produce
    representations invariant to noise conditions for ASR.  We evaluate the proposed architecture on the Aurora-4 task, a popular benchmark for
    noise robust ASR. We show that our method generalizes better than the standard multi-condition training especially when only a few noise categories are seen during training.
\end{abstract}
%
%\begin{keywords}
%speech recognition, deep neural networks, invariant representations, adversarial training
%\end{keywords}
%
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR)
    is the mismatch between the training and testing acoustic conditions. During
    testing, a system may encounter new recording conditions, microphone types, speakers,
    accents and types of background noises. Furthermore, even if the test scenarios are seen during training, there can be significant variability in their statistics. Thus, its important to develop ASR systems that are invariant to unseen acoustic conditions.
    %bhuv
    %Even in case all testing conditions have also appeared
    %in the training data, their acoustic statistics likely to be
    %different. For this reason, it's important to develop ASR systems that are
    %robust towards such undesirable acoustic variabilities.

    %While many types of model and feature based adaptation methods such as, maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
    %\cite{leggetter1995maximum} have been proposed, most of these methods were
    %designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
        %model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
    %methods Moreover, the discriminative gradient-based optimization used to train
    %DNNs, in combination with the ability of DNNs to learn highly non-linear
%feature transformations, allows for greater flexibility in
    %constructing training objective functions that promote learning of noise invariant
    %representations.
    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR 
    and iVectors~\cite{ivec} have been proposed to handle speaker variability; and Noise Adaptive Training (NAT)~\cite{msftnoise}  
    and Vector Taylor Series (VTS) to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
    acoustic models for ASR \cite{hinton2012deep, msft, ibm}, end-to-end systems are being proposed~\cite{essen, scatterTransforms, fromaudio} 
    for modeling the acoustic conditions within a single network. This allows us to take advantage of the network's ability to learn 
    highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning 
    of noise invariant representations. The main idea of this work is to force the acoustic model to learn a representation invariant to 
    noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}). This type of 
    noise-invariant training requires noise-condition labels during training only. It is related to the idea of generative adversarial 
    networks (GAN) and the gradient reverse method proposed in~\cite{goodfellow2014generative} and~\cite{ganin2014unsupervised} respectively 
    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

@@

    % bhuv :We should say how this is different form multi-task learning
    Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
    which can be adapted to new domains. The training data consists of the images
    labeled with classes of interest and separate domain (image background) labels. 
    %For example, the domain may refer to image background. The testing set does not need
   % to have the domain labels. 
    The network has a $Y$-like structure: the image is fed to the
    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
    a target classifier network (R). The goal of training is to learn the hidden 
    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
    the \emph{gadient reverse method} makes domain distributions similar to each other.

@@

Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method. It is a feed-forward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition). This branch is discarded in the testing phase. In our experiments we
used the noise condition as the domain label merging all noise types into one label
and clean as the other label. Our training loss function is  Eq.~\ref{eq:grm} with $\alpha$ set to 1 and
$L_3$ to $-d\log(1 - \hat{d}) - (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
of an incorrect domain classification in contrast to the Eq.~\ref{eq:grm} where the 
correct classification is minimized.
%{\bf Contrast this with term 3 of the original gradient reverse equation.}. 
The terms $L_1$ and $L_2$ are 
regular cross-entropies which are minimized with corresponding parameters $\theta_E$ and $\theta_D$.
For simplicity, we use only a single hyper-parameter -- the weight of the third term.

@@

\begin{table}
    \centering
    \caption{Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
        including seen and unseen noise and unseen microphone. First column
        is the number of noise conditions used for the training. The last row is a 
        preliminary experiment with layer-wise pretraining close to state-of-the-art
        model and a corresponding invariance training starting with a pretrained model.}
    \label{tab:results}
    \begin{tabular}{r|cc||cc|cc|cc|cc}
    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
    %           & wv1  & wv2  & wv1  & wv2 \\
        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
               & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
    \hline
    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
    \hline\hline
    6* &11.88        &11.99
    \end{tabular}
    %TODO: add numbers!
\end{table}

>>>>

\begin{document}
%\ninept
%
\maketitle
%
\begin{abstract}
    Modern automatic speech recognition (ASR) systems need to be robust under acoustic variability arising from environmental, speaker, channel, and recording conditions. Ensuring such robustness to variability is a challenge in modern day neural network-based ASR systems, especially when all types of variability are not seen during training. We attempt to address this problem by encouraging the neural network acoustic model to learn invariant feature representations.
    We use ideas from recent research on image generation using
    Generative Adversarial Networks and domain adaptation ideas extending
    adversarial gradient-based training. A recent work from Ganin et al. proposes to
    use adversarial training for image domain adaptation by using an intermediate
    representation from the main target classification network to deteriorate the domain 
    classifier performance through a separate neural network.
    Our work focuses on investigating neural architectures which produce
    representations invariant to noise conditions for ASR.  We evaluate the proposed architecture on the Aurora-4 task, a popular benchmark for
    noise robust ASR. We show that our method generalizes better than the standard multi-condition training especially when only a few noise categories are seen during training.
\end{abstract}
%
%\begin{keywords}
%speech recognition, deep neural networks, invariant representations, adversarial training
%\end{keywords}
%
\section{Introduction}
\label{sec:intro}
    One of the most challenging aspects of automatic speech recognition (ASR)
    is the mismatch between the training and testing acoustic conditions. During
    testing, a system may encounter new recording conditions, microphone types, speakers,
    accents and types of background noises. Furthermore, even if the test scenarios are seen during training, there can be significant variability in their statistics. Thus, its important to develop ASR systems that are invariant to unseen acoustic conditions.
    %bhuv
    %Even in case all testing conditions have also appeared
    %in the training data, their acoustic statistics likely to be
    %different. For this reason, it's important to develop ASR systems that are
    %robust towards such undesirable acoustic variabilities.

    %While many types of model and feature based adaptation methods such as, maximum likelihood linear regression (MLLR) and feature-space MLLR (fMLLR)
    %\cite{leggetter1995maximum} have been proposed, most of these methods were
    %designed with Gaussian mixture model-hidden Markov model (GMM-HMM) systems in mind. Since the enormous success of
        %model adaptation and feature-space adaptation are not optimal for the overall hybrid DNN-HMM system.
    %methods Moreover, the discriminative gradient-based optimization used to train
    %DNNs, in combination with the ability of DNNs to learn highly non-linear
%feature transformations, allows for greater flexibility in
    %constructing training objective functions that promote learning of noise invariant
    %representations.
    Several model and feature based adaptation methods such as Maximum Likelihood Linear Regression (MLLR),  feature-based MLLR 
    and iVectors~\citep{saon2013speaker} have been proposed to handle speaker variability; and Noise Adaptive Training \citep[NAT;][]{kalinli2010noise}
    and Vector Taylor Series \citep[VTS;][]{un1998speech} to handle environment variability. With the increasing success of Deep Neural Network (DNN) 
    acoustic models for ASR \citep{hinton2012deep,seide2011conversational,sainath2011making}, end-to-end systems are being proposed~\citep{miao2015eesen,sainath2015learning} 
    for modeling the acoustic conditions within a single network. This allows us to take advantage of the network's ability to learn 
    highly non-linear feature transformations, with greater flexibility in constructing training objective functions that promote learning 
    of noise invariant representations. The main idea of this work is to force the acoustic model to learn a representation invariant to 
    noise conditions, instead of explicitly using noise robust acoustic features (Section~\ref{sec:invariant-speech}). This type of 
    noise-invariant training requires noise-condition labels during training only. It is related to the idea of generative adversarial 
    networks (GAN) and the gradient reverse method proposed in~\cite{goodfellow2014generative} and~\cite{ganin2014unsupervised} respectively 
    (Section~\ref{sec:relatedwork}). We present results on the Aurora-4 speech recognition task in Section~\ref{sec:experiments} and summarize 
    our findings in Section~\ref{sec:discussion}.

@@

    % bhuv :We should say how this is different form multi-task learning
    Prior work by~\cite{ganin2014unsupervised} proposed a method of training a network 
    which can be adapted to new domains. The training data consists of the images
    labeled with classes of interest and separate domain (image background) labels. 
    %For example, the domain may refer to image background. The testing set does not need
   % to have the domain labels. 
    The network has a $Y$-like structure: the image is fed to the
    first network which produces a hidden representation $h$. Then this representation $h$ is input to two separate networks: a domain classifier network (D) and 
    a target classifier network (R). The goal of training is to learn the hidden 
    representation that is invariant to the domain labels and performs well on the target classification task, so that the domain information doesn't 
    interfere with the target classifier at test time. Similar to the GAN objective, which forces the generation distribution be close to the data distribution,
    the \emph{gradient reverse method} makes domain distributions similar to each other.

@@

Figure~\ref{fig:model} depicts the model, which is same as the model for the gradient reverse method. It is a feed-forward neural network trained to predict the CD HMM state, with a branch that predicts the domain (noise condition). This branch is discarded in the testing phase. In our experiments we
used the noise condition as the domain label merging all noise types into one label
and clean as the other label. Our training loss function is  Eq.~\ref{eq:grm} with $\alpha$ set to 1 and
$L_3$ to $d\log(1 - \hat{d}) + (1-d)\log(\hat{d})$. $L_3$ term maximizes the probability
of an incorrect domain classification in contrast to the gradient reverse where the 
correct classification is minimized.
%{\bf Contrast this with term 3 of the original gradient reverse equation.}. 
The terms $L_1$ and $L_2$ are 
regular cross-entropies which are minimized with corresponding parameters $\theta_E$ and $\theta_D$.
For simplicity, we use only a single hyper-parameter -- the weight of the third term.

@@

\begin{table}
    \centering
    \caption{Average word error rate (WER\%) on Aurora-4 dataset on all test conditions,
        including seen and unseen noise and unseen microphone. First column
        is the number of noise conditions used for the training. The last row is a 
        preliminary experiment with layer-wise pre-training close to state-of-the-art
        model and a corresponding invariance training starting with a pretrained model.}
    \label{tab:results}
    \begin{tabular}{r|cc||cc|cc|cc|cc}
    %\multirow{2}{*}{Subset} & \multicolumn{2}{c}{Orig}  & \multicolumn{2}{|c}{Inv}\\ 
    %           & wv1  & wv2  & wv1  & wv2 \\
        Noise       &Inv&BL&  \multicolumn{2}{c|}{A} & \multicolumn{2}{c|}{B} & \multicolumn{2}{c|}{C} & \multicolumn{2}{c}{D}\\
               & & &  Inv & BL & Inv & BL & Inv & BL & Inv & BL\\
    \hline
    1           &16.36        &18.14 &6.54&7.57    &12.71& 14.09   & 11.45&   13.10    & 22.47 &   24.80    \\
    2           &15.56        &17.39 &5.90&  6.58 &   11.69   &13.28   &11.12   &13.51   &21.79   &23.96 \\
    3           &14.24        &14.67 &5.45 & 5.08&    10.76&   12.44&   9.75&    9.84 &   19.93&   19.30\\
    4           &13.61        &13.84 & 5.08 &5.29    &9.73    &9.97    &9.49    &9.56    &19.49   &19.90\\         
    5           &13.41        &13.02 & 5.12 &5.34    &9.52    &9.42    &9.55    &8.67    &19.33   &18.65\\         
    6           &12.62        &12.60 & 4.80 &4.61    &9.04    &8.86    &8.76    &8.59    &18.16   &18.21\\
    \hline\hline
    6* &11.88        &11.99
    \end{tabular}
    %TODO: add numbers!
\end{table}
